---
title: "Combine_outputs_sens"
author: 'Noémie Huvelle'
date: "2025-12-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Prepare R environment 
```{r load packages}
rm(list = ls())
library(ggplot2)
library(dplyr)
library(tidyr)
library(rlang)
library(here)
```

```{r set directories}
wd <- here("run_scenarios/QualityFlo_treatment")
in_files <- here("run_scenarios/QualityFlo_treatment/Intermediate_results")
out_file <- here("run_scenarios/QualityFlo_scenarios")
setwd(wd)
```

```{r load sensitivity analysis data}

files <- list.files(in_files, pattern = "\\.csv$", full.names = TRUE)

sens_data_list <- setNames(lapply(files, read.csv),
                            basename(files))

sens_data_list <- lapply(sens_data_list, function(df) {
  names(df)[1] <- "doy"
  df
})

str(sens_data_list)
```

```{r load baseline data}
base_flo_ressources <- read.csv(paste(out_file,
                                      "/baseline_ResFlo.csv",
                                      sep=""))
str(base_flo_ressources)
```

#Create scenario dataframes and save them 
```{r function to properly combine df}
prepare_sens_df <- function(sens_df,
                            base_df,
                            join_by   = "doy",
                            join_col  = "Prop_foraging_day",
                            remove_cols = "date",
                            rename_cols = c(
                              Quality_crop        = "quality_crop",
                              Quality_nat         = "quality_nat",
                              Prop_foraging_crop  = "prop_forag_crop",
                              Prop_foraging_day   = "prop_forag_day"
                            )) {

  ## --- 1. JOIN ---
  sens_df <- sens_df %>%
    dplyr::left_join(
      base_df %>% dplyr::select(dplyr::all_of(join_by), dplyr::all_of(join_col)),
      by = join_by
    )

  ## --- 2. REMOVE columns if present ---
  if (!is.null(remove_cols)) {
    sens_df <- sens_df %>% dplyr::select(-dplyr::all_of(intersect(remove_cols, names(sens_df))))
  }

  ## --- 3. SAFE RENAME (only rename if column exists!)
  for (new_name in names(rename_cols)) {
    old_name <- rename_cols[[new_name]]
    if (old_name %in% names(sens_df)) {
      names(sens_df)[names(sens_df) == old_name] <- new_name
    }
  }

  ## --- 4. REORDER columns ---
  final_order <- c("doy", "Prop_foraging_day",
                   "Quality_crop", "Quality_nat",
                   "Prop_foraging_crop")

  final_order <- intersect(final_order, names(sens_df))

  sens_df <- sens_df %>%
    dplyr::select(dplyr::all_of(final_order), dplyr::everything())

  return(sens_df)
}

```

## Quality crop sensitivity
```{r quality crop sensitivity df}
sens_crop_low <- prepare_sens_df(
  sens_df   = sens_data_list$scenario_QF_QC_low.csv,
  base_df   = base_flo_ressources,
  join_by   = "doy",
  join_col  = "Prop_foraging_day",
  remove_cols = "date"
)
str(sens_crop_low) #OK <3
write.csv(sens_crop_low,
          paste(out_file,
                "/scenario_QC_low.csv",
                sep=""),
          row.names = FALSE)

sens_crop_high <- prepare_sens_df(
  sens_df   =  sens_data_list$scenario_QF_QC_high.csv,
  base_df   = base_flo_ressources,
  join_by   = "doy",
  join_col  = "Prop_foraging_day",
  remove_cols = "date"
)
str(sens_crop_high)
write.csv(sens_crop_high,
          paste(out_file,
                "/scenario_QC_high.csv",
                sep=""),
          row.names = FALSE)

sens_crop_high$Quality_crop == sens_crop_low$Quality_crop
```

## Quality nat sensitivity
```{r quality nat sensitivity df}
sens_nat_low <- prepare_sens_df(
  sens_df   = sens_data_list$scenario_QF_QN_low.csv,
  base_df   = base_flo_ressources,
  join_by   = "doy",
  join_col  = "Prop_foraging_day",
  remove_cols = "date"
)
str(sens_nat_low)
write.csv(sens_nat_low,
          paste(out_file,
                "/scenario_QN_low.csv",
                sep=""),
          row.names = FALSE)

sens_nat_high <- prepare_sens_df(
  sens_df   = sens_data_list$scenario_QF_QN_high.csv,
  base_df   = base_flo_ressources,
  join_by   = "doy",
  join_col  = "Prop_foraging_day",
  remove_cols = "date"
)
str(sens_nat_high)
write.csv(sens_nat_high,
          paste(out_file,
                "/scenario_QN_high.csv",
                sep=""),
          row.names = FALSE)


sens_nat_high$Quality_nat == sens_nat_low$Quality_nat
```
There is an issue here... The quality_nat shouldnt be the same AND the issue stems from the Sens_analysis/Scenario_sensitivity_nat_lowORhigh -> ISSUE SOLVED

## Prop foraging crop sensitivity
Sensitivity of the alpha parameter
```{r prop foraging crop sensitivity df}
sens_prop_forag_crop_low <- prepare_sens_df(
  sens_df   = sens_data_list$scenario_QF_PFC_low.csv,
  base_df   = base_flo_ressources,
  join_by   = "doy",
  join_col  = "Prop_foraging_day",
  remove_cols = "date"
)
str(sens_prop_forag_crop_low)
write.csv(sens_prop_forag_crop_low,
          paste(out_file,
                "/scenario_PFC_low.csv",
                sep=""),
          row.names = FALSE)

sens_prop_forag_crop_high <- prepare_sens_df(
  sens_df   = sens_data_list$scenario_QF_PFC_high.csv,
  base_df   = base_flo_ressources,
  join_by   = "doy",
  join_col  = "Prop_foraging_day",
  remove_cols = "date"
)
str(sens_prop_forag_crop_high)
write.csv(sens_prop_forag_crop_high,
          paste(out_file,
                "/scenario_PFC_high.csv",
                sep=""),
          row.names = FALSE)
```

## Weather thresholds sensitivity
### Precipitation thresholds
```{r PFD - Precip sensitivity df}
sens_Precip_low <- prepare_sens_df(
  sens_df   = sens_data_list$scenario_PFD_Precip_low.csv,
  base_df   = base_flo_ressources,
  join_by   = "doy",
  join_col  = c("Quality_crop", "Quality_nat", "Prop_foraging_crop"),
  remove_cols = "date"
)
str(sens_Precip_low)
write.csv(sens_Precip_low,
          paste(out_file,
                "/scenario_PFD_Precip_low.csv",
                sep=""),
          row.names = FALSE)


sens_Precip_high <- prepare_sens_df(
  sens_df   = sens_data_list$scenario_PFD_Precip_high.csv,
  base_df   = base_flo_ressources,
  join_by   = "doy",
  join_col  = c("Quality_crop", "Quality_nat", "Prop_foraging_crop"),
  remove_cols = "date"
)
str(sens_Precip_high)
write.csv(sens_Precip_high,
          paste(out_file,
                "/scenario_PFD_Precip_high.csv",
                sep=""),
          row.names = FALSE)

sens_Precip_high$Prop_foraging_day == sens_Precip_low$Prop_foraging_day
```

### Temperature thresholds
```{r  PFD - Temperature sensitivity df}
sens_T_low <- prepare_sens_df(
  sens_df   = sens_data_list$scenario_PFD_T_low.csv,
  base_df   = base_flo_ressources,
  join_by   = "doy",
  join_col  = c("Quality_crop", "Quality_nat", "Prop_foraging_crop"),
  remove_cols = "date"
)
str(sens_T_low)
write.csv(sens_T_low,
          paste(out_file,
                "/scenario_PFD_T_low.csv",
                sep=""),
          row.names = FALSE)

sens_T_high <- prepare_sens_df(
  sens_df   = sens_data_list$scenario_PFD_T_high.csv,
  base_df   = base_flo_ressources,
  join_by   = "doy",
  join_col  = c("Quality_crop", "Quality_nat", "Prop_foraging_crop"),
  remove_cols = "date"
)
str(sens_T_high)
write.csv(sens_T_high,
          paste(out_file,
                "/scenario_PFD_T_high.csv",
                sep=""),
          row.names = FALSE)
```

### Relative Humidity thresholds
```{r PFD - RH sensitivity df}
sens_RH_low <- prepare_sens_df(
  sens_df   = sens_data_list$scenario_PFD_RH_low.csv,
  base_df   = base_flo_ressources,
  join_by   = "doy",
  join_col  = c("Quality_crop", "Quality_nat", "Prop_foraging_crop"),
  remove_cols = "date"
)
str(sens_RH_low)
write.csv(sens_RH_low,
          paste(out_file,
                "/scenario_PFD_RH_low.csv",
                sep=""),
          row.names = FALSE)

sens_RH_high <- prepare_sens_df(
  sens_df   = sens_data_list$scenario_PFD_RH_high.csv,
  base_df   = base_flo_ressources,
  join_by   = "doy",
  join_col  = c("Quality_crop", "Quality_nat", "Prop_foraging_crop"),
  remove_cols = "date"
)
str(sens_RH_high)
write.csv(sens_RH_high,
          paste(out_file,
                "/scenario_PFD_RH_high.csv",
                sep=""),
          row.names = FALSE)
```
